Boosting, Sparsity, and Conditional Gradient: A Tutorial

### 2014-03-27 arXiv/sparsity-result.tex
fixed some small errors

% define strong convexity
The convergence rate of Algorithm~\ref{alg:comid} depends on the degree of 
convexity of the functions $\psi$ and $h_t$. We can measure this convexity 
using any norm $\|\cdot\|$ on $\W$.
\begin{definition}
A differentiable function $f$ is said to be \emph{$\rho$-strongly convex} with 
respect to a norm $\|\cdot\|$ if
\[ f(w) - f(w') - \nabla f(w')^{\top}(w - w') \geq \frac{\rho}{2}\|w-w'\|^2 \]
for all $w, w' \in \W$.
\end{definition}
When $\rho = 1$ we sometimes just say that $f$ is strongly convex with respect to $\|\cdot\|$.
If $\|\cdot\|$ is the $l^2$ norm and $f$ is twice-differentiable, then $f$ is $\rho$-strongly 
convex if and only if $\nabla^2 f(w) \succeq \rho I$ for all $w \in \W$. In particular, 
$f(w) = \frac{1}{2}\|w - w_0\|_2^2$ is $1$-strongly convex with respsect to $\|\cdot\|_2$ for 
any $w_0$, although this is not true for general norms (for instance, 
$\frac{1}{2}\|w-w_0\|_1^2$ is not strongly convex with respect to $\|\cdot\|_1$ for any value 
of $\rho$). See Section~\ref{sec:fenchel} for more details.

@@

\begin{corollary}
Suppose that $\psi$ is $\rho_0$-strongly convex with respect to $\|\cdot\|$, but make 
no assumptions on $h_t$ other than convexity. Then, for $\alpha_t = 1/\sqrt{T}$, we have
\[ f(w^*) \leq \inf_w \left\{ f(w) + \frac{1}{\sqrt{T}} \psi(w)\right\} + \frac{r^2}{2\rho_0\sqrt{T}}. \]
\end{corollary}
\begin{proof}
We have $A_T = \sqrt{T}$ and $R_t = \rho_0$. Then Theorem~\ref{thm:comid} yields
\begin{align*}
f(w^*) &\leq \inf_w \left\{ f(w) + \frac{1}{\sqrt{T}} \psi(w)\right\} + \frac{r^2}{2\sqrt{T}} \sum_{t=1}^T \frac{1}{\rho_0 T} \\
 &= \inf_w \left\{ f(w) + \frac{1}{\sqrt{T}} \psi(w)\right\} + \frac{r^2}{2\rho_0\sqrt{T}}.
\end{align*}
\end{proof}
So, if the $h_t$ are strongly convex, we can obtain an $O\left(\frac{1}{T}\right)$ 
convergence rate. If the $h_t$ are only (weakly) convex, we can still obtain an 
$O\left(\frac{1}{\sqrt{T}}\right)$ convergence rate, by ``deforming'' $f$ by a 
strongly convex function $\psi$.

@@

This leads to the following primal-dual algorithm:
\begin{algorithm}
\caption{Primal-dual Composite Mirror Descent}
\label{alg:boosted-comid}
\begin{algorithmic}
\STATE Initialize $\theta_0 \gets 0$, $\psi_0 \gets \psi$
\FOR{$t = 1$ {\bfseries to} $T$}
  \STATE $\psi_t \gets \psi_{t-1} + \alpha_t h$
  \STATE $w_t \gets \argmin_{w \in \W} \psi_t(w) + \theta_{t-1}^{\top} w$
  \STATE $v_t \gets \argmax_{v \in S} f(v, w_t)$
  \STATE $\theta_t \gets \theta_t + \alpha_t \left(\frac{\partial f(v_t, w_t)}{\partial w_t} - \frac{\partial h(w_t)}{\partial w_t}\right)$
\ENDFOR
\STATE Output $(v^*, w^*) = \left(\frac{\sum_{t=1}^T \alpha_t v_t}{\sum_{t=1}^T \alpha_t}, \frac{\sum_{t=1}^T \alpha_t w_t}{\sum_{t=1}^T \alpha_t}\right)$
\end{algorithmic}
\end{algorithm}

@@

\begin{corollary}
Suppose that $\rho_0 > 0$ and set $\alpha_t = 1/\sqrt{T}$. Then
\[ \Delta_T \leq \frac{1}{\sqrt{T}} \left[ \psi(w_T) + \frac{r^2}{2\rho_0} \right]. \]
\end{corollary}

>>>>

% define strong convexity
The convergence rate of Algorithm~\ref{alg:comid} depends on the degree of 
convexity of the functions $\psi$ and $h_t$. We can measure this convexity 
using any norm $\|\cdot\|$ on $\W$.
\begin{definition}
A differentiable function $f$ is said to be \emph{$\rho$-strongly convex} with 
respect to a norm $\|\cdot\|$ if
\[ f(w) - f(w') - \nabla f(w')^{\top}(w - w') \geq \frac{\rho}{2}\|w-w'\|^2 \]
for all $w, w' \in \W$.
\end{definition}
When $\rho = 1$ we sometimes just say that $f$ is strongly convex with respect to $\|\cdot\|$.
If $\|\cdot\|$ is the $l^2$ norm and $f$ is twice-differentiable, then $f$ is $\rho$-strongly 
convex if and only if $\nabla^2 f(w) \succeq \rho I$ for all $w \in \W$. In particular, 
$f(w) = \frac{1}{2}\|w - w_0\|_2^2$ is $1$-strongly convex with respect to $\|\cdot\|_2$ for 
any $w_0$, although this is not true for general norms (for instance, 
$\frac{1}{2}\|w-w_0\|_1^2$ is not strongly convex with respect to $\|\cdot\|_1$ for any value 
of $\rho$). See Section~\ref{sec:fenchel} for more details.

@@

\begin{corollary}
Suppose that $\psi$ is $\rho_0$-strongly convex with respect to $\|\cdot\|$, but make 
no assumptions on $h_t$ other than convexity. Then, for $\alpha_t = 2/\sqrt{T}$,
\fTBD{JHH: corrected choice of $\alpha_{t}$ and statement of theorem; also added in a necessary log factor} we have
\[ f(w^*) \leq \inf_w \left\{ f(w) + \frac{1}{\sqrt{T}} \psi(w)\right\} + \frac{2r^2 \log(T+1)}{\rho_0\sqrt{T}}. \]
\end{corollary}
\begin{proof}
We have $A_T \ge \sqrt{T}$ and $R_t = \rho_0$ so Theorem~\ref{thm:comid} yields
\begin{align*}
f(w^*) 
&\leq \inf_w \left\{ f(w) + \frac{1}{\sqrt{T}} \psi(w)\right\} + \frac{r^2}{2\sqrt{T}} \sum_{t=1}^T \frac{4}{\rho_0 t} \\
&\leq \inf_w \left\{ f(w) + \frac{1}{\sqrt{T}} \psi(w)\right\} + \frac{2r^2\log(T+1)}{\rho_0\sqrt{T}}.
\end{align*}
\end{proof}
So, if the $h_t$ are strongly convex, we can obtain an $O\left(\frac{1}{T}\right)$ 
convergence rate. If the $h_t$ are only (weakly) convex, we can still obtain an 
$O\left(\frac{\log(T+1)}{\sqrt{T}}\right)$ convergence rate, by ``deforming'' $f$ by a 
strongly convex function $\psi$.

@@

This leads to the following primal-dual algorithm:
\begin{algorithm}
\caption{Primal-dual Composite Mirror Descent}
\label{alg:boosted-comid}
\begin{algorithmic}
\STATE Initialize $\theta_0 \gets 0$, $\psi_0 \gets \psi$
\FOR{$t = 1$ {\bfseries to} $T$}
  \STATE $\psi_t \gets \psi_{t-1} + \alpha_t h$
  \STATE $w_t \gets \argmin_{w \in \W} \psi_t(w) + \theta_{t-1}^{\top} w$
  \STATE $v_t \gets \argmax_{v \in S} f(v, w_t)$
  \STATE $\theta_t \gets \theta_{t-1} + \alpha_t \left(\frac{\partial f(v_t, w_t)}{\partial w_t} - \frac{\partial h(w_t)}{\partial w_t}\right)$
\ENDFOR
\STATE Output $(v^*, w^*) = \left(\frac{\sum_{t=1}^T \alpha_t v_t}{\sum_{t=1}^T \alpha_t}, \frac{\sum_{t=1}^T \alpha_t w_t}{\sum_{t=1}^T \alpha_t}\right)$
\end{algorithmic}
\end{algorithm}

@@

\begin{corollary}
Suppose that $\rho_0 > 0$ and set $\alpha_t = 2/\sqrt{T}$. Then
\[ \Delta_T \leq \frac{1}{\sqrt{T}} \left[ \psi(w_T) + \frac{2r^2\log(T+1)}{\rho_0} \right]. \]
\end{corollary}
